---
title: "GB_FinalProject"
author: "GB"
date: "2026-04-21"
output: html_document
---

# Loading all required libraries and set chunk options
```{r setup, include=FALSE}
library(ggplot2)
library(tidyverse)
library(lme4)
library(emmeans)
library(lmerTest)
library(dplyr)
library(multcompView)
library(multcomp)
library(fs)
```


# Loading the dataset

```{r}
mortality <- read.csv("Mortality.csv")
head(mortality)
```
# Calculate percent mortality for each observation
$$ ((InitialJ2 - FinalJ2) / InitialJ2) * 100 $$


```{r}
Mortality <- mortality %>% 
   mutate(Mortality_percent = ((initialJ2 - finalJ2) / initialJ2) * 100)
head(Mortality)
```


# Summary Statistics by Treatment
```{r}
summary_stats <- Mortality %>%
  group_by(Treatment) %>% # Group by treatment and calculate mean, SD, SE
  summarise(n = n(),
    mean_mortality = mean(Mortality_percent),
            sd_mortality = sd(Mortality_percent),
            se = sd_mortality/n )

# # Display as table

knitr::kable(summary_stats)

```

# selecting Top Treatments with Mean mortality higher than 50% and untreated control

```{r}
# Filter strains with >50% mortality or named 'UntreatedControl'
selected_strains <- summary_stats %>%
  filter(mean_mortality > 25 | Treatment == "UntreatedControl")
selected_strains

```


# Plot Manuscript-Ready Figure
```{r}
# Barplot of selected treatments, color-coded with red for Untreatedcontrol and green for Fluopyram
ggplot(selected_strains, aes(x = reorder(Treatment, -mean_mortality), y = mean_mortality, color="black")) +
  geom_col(aes(fill = Treatment == "UntreatedControl")) +
   scale_fill_manual(values = c("TRUE" = "red", "FALSE" = "blue")) +
  geom_col(aes(fill = Treatment == "Fluopyram")) +
  scale_fill_manual(values = c("TRUE" = "Black", "FALSE" = "blue")) +
  geom_text(aes(label = round(mean_mortality, 1)), hjust = -0.1, size = 3) +
  coord_flip() +
  ylab(" Mortality (%)") +
  xlab("PGPR Strain") +
  ggtitle("PGPR Strains with >25% Mortality + UntreatedControl") +
   theme_minimal(base_size = 10) +
  theme(plot.title = element_text(hjust = 0.5),
        legend.position = "none")

```




# Mixed Effects Model (GLMM)

```{r}
# Convert grouping variables to factors
Mortality$Treatment <- as.factor(Mortality$Treatment)
Mortality$Replication <- as.factor(Mortality$Replication)
```


```{r}
# Linear mixed model with rep as random effect
mod <- lmer(Mortality_percent  ~ Treatment + (1|Replication),
             data = Mortality)
# Summary of the model, includes fixed effects and random structure
summary(mod)
 
```


# ANOVA and LSMeans
```{r}
# ANOVA model
anova_mod <- aov(Mortality_percent ~ Treatment + Replication, data = Mortality)
summary(anova_mod)

# Calculate estimated marginal means (LSMeans) with Benferroni adjustment
lsm <- emmeans(anova_mod, pairwise ~ Treatment, adjust = "bon")
```





# Treatment vs Fluopyram Comparison
```{r}
Fluopyram <- contrast(lsm, "trt.vs.ctrl", ref = 1, adjust = "bon") # Compare each treatment to a reference (Fluopyram = 1)
Fluopyram 

```

# Treatment vs Untreated Control comparison
```{r}
untreatedcontrol <- contrast(lsm, "trt.vs.ctrl", ref = 217, adjust = "bon") # Compare each treatment to UntreatedControl (Untreated Control = 217)
untreatedcontrol
```



