---
title: "CodingChallenge6_IterationsandFunctions"
author: "GB"
date: "2026-03-19"
output: html_document
---

#Question1
1. Regarding reproducibility, what is the main point of writing your own functions
and iterations?

Ans: The main purpose is to make analyses consistent, repeatable, and less error-prone. When you write your own functions and loops:

The same steps are applied exactly the same way every time

You avoid manual repetition, which reduces human error

Your workflow becomes transparent and easy to reuse

Others (or your future self) can understand and reproduce the analysis by running the same code




#Question 2.  

In your own words, describe: 

(1)
how to write a function in R 
Ans: A function in R is a reusable block of code designed to perform a specific task.
- Basic syntax: We define a function  by   assigning  it  a name and using the function() keyword, with inputs(arguments) inside parentheses. 
-The actual operations are written inside curly braces{}
-The function returns the final results either explicitly using a return statement or implicitly as the last evaluated expression. 

Simply, we give the function inputs, it performs operations, and it oitputs a result. 

(2)
how to write a for loop in R. In your answer, be specific about: basic syntax, where the
code is written, and how results are returned or stored. (Answer in plain text; no code
required for this question.)

Ans: A for loop is used to repeat a set of operations over a sequence of values.

Basic syntax: It starts with 'for', followed by a variable that takes on each value in a sequence.

Where code is written: The repeated operations are placed inside curly braces {}.

Storing results: Results are usually stored in a pre-created object (like a vector or list), which gets updated during each iteration.

In simple terms, a for loop runs the same code multiple times, changing the input each time and storing the results.

#Question 3

```{r}
Cities <- read.csv(file.choose())
head(Cities)
```
```{r}
Haversine<- function(lat1, lon1, lat2, lon2) {rad.lat1 <- lat1 * pi/180
  rad.lon1 <- lon1 * pi/180
  rad.lat2 <- lat2 * pi/180
  rad.lon2 <- lon2 * pi/180
  
  # Haversine formula
  delta_lat <- rad.lat2 - rad.lat1
  delta_lon <- rad.lon2 - rad.lon1
  a <- sin(delta_lat / 2)^2 + cos(rad.lat1) * cos(rad.lat2) * sin(delta_lon / 2)^2
  c <- 2 * asin(sqrt(a))
  
  # Earth's radius in kilometers
  earth_radius <- 6378137
  
  # Calculate the distance
  distance_km <- (earth_radius * c) / 1000
  
  return(distance_km) 
}
colnames(Cities)




```
# Question 5
```{r}
auburn <- Cities$lat[Cities$city == "Auburn" ]
auburn2 <- Cities$long[Cities$city == "Auburn" ]

nyc <- Cities$lat[Cities$city == "New York"]
nyc2 <- Cities$long[Cities$city == "New York"]

# Calculate distance 
 auburn_nyc_dist <- Haversine(auburn, auburn2, nyc, nyc2)

auburn_nyc_dist
```
# Question 6 (Loop)
```{r}
for (i in 1:nrow(Cities)) {
  
  dist <- Haversine(auburn, auburn2,
                    Cities$lat[i],
                    Cities$long[i])
  
  print(dist)
}
```

# Qusetion 7 (dataframe)
```{r}
distance_df <- NULL

for (i in 1:nrow(Cities)) {
    result_i <- data.frame(
      City1 = Cities$city[i],
      City2 = "Auburn",
      Distance_km = Haversine(auburn, auburn2,
                              Cities$lat[i],
                              Cities$long[i]))
    
    distance_df <- rbind.data.frame(distance_df, result_i)
}
print(distance_df)
```



