---
title: "CodingChallenge4"
author: "GB"
date: "2026-02-21"
output:
  html_document:
    toc: true
    toc_float: true
    word_document: null
  pdf_document: null
  md_document:
    variant: gfm
  word_document:
    toc: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(ggplot2)
library(markdown)

```





```{r}
library(knitr)
library(ggplot2)
library(markdown)
library(tinytex)
```

#Link to my paper

- [Link to my paper]("C:\Users\gzb0078\OneDrive - Auburn University\Documents\Reproducibility-GB\papers\fungus_paper.pdf")


- [Paper](papers/fungus_paper.pdf)



```{r}
mycotoxin<- read.csv(file.choose(), na.strings = "na")
head(mycotoxin)

library(ggplot2)
cbbPalette <- c("#56B4E9", "#009E73")

library(ggrepel)

```

#Ques5
##DON_PLOT1

```{r}

Plot1 <- ggplot(mycotoxin, aes(x = Treatment, y = DON, fill = Cultivar)) + 
  geom_boxplot() + # add boxplot layer 
  xlab("") +  
  ylab("DON (ppm)") + # y label 
  geom_point(alpha = 0.6,pch = 21, color = "black", position = position_jitterdodge()) + 
  scale_fill_manual(values = cbbPalette)+ # transparency of the jittered points to 0.6. #Jitter points over the boxplot and fill the points and boxplots Cultivar with two colors from the cbbPallete  
  facet_wrap(~Cultivar)+ #faceted by Cultivar 
  theme_classic() # for classic theme  

Plot1
```

##15ADON
```{r}
Plot2 <- ggplot(mycotoxin, aes(x=Treatment, y=X15ADON, fill= Cultivar))+
  geom_boxplot()+
  xlab("")+
  ylab("15ADON")+
  geom_point(alpha=0.6,pch=21, color="black", position= position_jitterdodge())+
  facet_wrap(~Cultivar)+
  theme_classic()

Plot2
```

##MassperSeed_mg_PLOT3

```{r}

Plot3 <- ggplot(mycotoxin, aes(x = Treatment, y = MassperSeed_mg, fill = Cultivar)) +
  geom_boxplot() + # add boxplot layer
  xlab("") +  
  ylab("Seed Mass (mg)") + # y label
  geom_point(alpha = 0.6,pch = 21, color = "black", position = position_jitterdodge()) +
  scale_fill_manual(values = cbbPalette)+ # transparency of the jittered points to 0.6. #Jitter points over the boxplot and fill the points and boxplots Cultivar with two colors from the cbbPallete 
  facet_wrap(~Cultivar)+ #faceted by Cultivar
  theme_classic() # for classic theme 

Plot3
```
##3Combined_PLOT

```{r}

install.packages("ggpubr")

library(ggpubr)

figure=ggarrange(
  Plot1,
  Plot2,
  Plot3,
  labels = c("A", "B", "C"),
  nrow = 1,
  ncol = 3,
  common.legend=TRUE
)

figure
```

### Plot 1: plot1 with pairwise comparisons

```{r}
 
Plot1_pwc <- Plot1 +
  geom_pwc(aes(group = Treatment), method = "t.test", label = "p.adj.format") +  # Add pairwise t-tests
  theme_classic()

Plot1_pwc
```
###Plot 2: plot2 with pairwise comparisons

```{r}

Plot2_pwc <- Plot2 +
  geom_pwc(aes(group=Treatment), method= "t.test", label= "p.adj.format")+
  theme_classic()

Plot2_pwc
```
###Plot 3: plot3 with pairwise comparisons

```{r}
Plot3_pwc <- Plot3 + geom_pwc(aes(group=Treatment), method = "t.test", label = "p.adj.format")+
  theme_classic()
Plot3_pwc

```
##Combine all three plots with a common legend
```{r}

Combined_pwc =ggarrange(
  Plot1_pwc,
  Plot2_pwc,
  Plot3_pwc,
  labels = c("D", "E", "F"),
  nrow = 1,
  ncol = 3,
  common.legend=TRUE
)

Combined_pwc
```



