---
title: "CodingChallenge5"
author: "GB"
date: "2026-03-19"
output:
  word_document:
     toc: true 
  md_document:
    variant: gfm
  html_document:
    df_print: paged
  pdf_document: default
---
# Question 1
```{r}
DiversityData <-read.csv(file.choose())
Metadata<- read.csv(file.choose())
head(DiversityData)
head(Metadata)
```
# Question 2
Join the two dataframes together by the common column ‘Code’.
```{r}
alpha <- merge(DiversityData, Metadata, by = "Code")
colnames(alpha)
head(alpha)
```
# Question 3
Calculate Pielou’s evenness index: Pielou’s evenness is an ecological parameter calculated by
the Shannon diversity index (column Shannon) divided by the log of the richness column.

Using mutate, create a new column to calculate Pielou’s evenness index.
```{r}
library(dplyr)
alpha_even <- alpha %>%
  mutate(Pielou_evenness = shannon / log(richness))
colnames(alpha_even)
```
# Question 4
Using tidyverse language of functions and the pipe, use the summarise function and tell me
the mean and standard error evenness grouped by crop over time.
```{r}
library(tidyverse)
alpha_average <- alpha_even %>%
  group_by(Crop, Time_Point) %>%
  summarise(
    Mean.even = mean(Pielou_evenness),
    n = n(),
    sd.dev = sd(Pielou_evenness)
  ) %>%
  mutate(std.err = sd.dev / sqrt(n))
```
# Question 5
Select relevant columns: select the columns Time_Point, Crop, and mean.even.
Reshape the data: Use the pivot_wider function to transform the data from long to wide format, creating new columns for each Crop with values from mean.even.
Calculate the difference between the soybean column, the soil column, and the difference between the cotton column and the soil column
```{r}
alpha_average2 <- alpha_average %>%
  select(Time_Point, Crop, Mean.even) %>%
  pivot_wider(names_from = Crop, values_from = Mean.even) %>%
  mutate(
    diff.cotton.even = Soil - Cotton,
    diff.soybean.even = Soil - Soybean
  )

head(alpha_average2)
```

# Question 6

Select relevant columns: select the columns Time_Point, diff.cotton.even, and diff.soybean.even.
Reshape the data: Use the pivot_longer function to transform the data from wide to long format, creating a new column named diff that contains the values from diff.cotton.even and diff.soybean.even.
```{r}
plot_data <- alpha_average2 %>%
  select(Time_Point, diff.cotton.even, diff.soybean.even) %>%
  pivot_longer(
    c(diff.cotton.even, diff.soybean.even),
    names_to = "diff"
  )
```

```{r}
ggplot(plot_data, aes(x = Time_Point, y = value, color = diff)) +
  geom_line() +
  theme_minimal(base_size = 8, base_family = "Arial") + theme(panel.grid = element_blank(), axis.line = element_line(colour = "Black")) + labs(x = "Time (hrs)",
    y = "Difference from soil in Pielou's evenness",
    color = "diff"
  )
```

# Question 7
[Coding Challenge 5](https://github.com/GayatriBhandari/Reproducibility-GB.git)

